Accelerating structure-function mapping using the ViVa webtool to mine natural variation

### 2019-09-26 manuscript/02-results.Rmd
final submission, minus a few copyedits

(ref:app-workflow) **ViVa workflow** (a) Workflow diagram of ViVa. Blue indicates user actions, yellow indicates processing steps performed by the application. (b) Detailed look at mapping tab parameters: User selects which genes to look at (1) then clicks the Submit button (2). The "Allele selection" pane is then filled in with all non-reference variants meeting the criteria. The user can adjust the range of nucleotide diversity and the type of SNP with a slider and radio buttons (3) to make the list of variants a manageable size. The list of variants is updated as changes are made to these controls. The user then selects variants to display on the map (4). Clicking the Update Map button (5) populates the map below with points located at the collection coordinates of each accession and colored by the selected alleles. (c) Detailed look at the browser tab options: The user first selects which genes to analyze and clicks submit (1). The “hide 0|0 genotype?” checkbox (2) removes rows from the table containing the reference allele. Four configurable filters of two types are provided. The first filter type is text matching (3); the user selects a column to filter from a drop-down menu then enters one or more text strings to match in that column. The second type of filter is numeric range matching (4); the user again selects a column, then specifies a range to match by typing or selecting minimum and maximum values. After configuring the filters, the user clicks the Apply Filters button (5) to update the table. The table is updated such that only rows that meet all filter conditions are displayed.
```{r app-workflow, fig.cap = "(ref:app-workflow)"}
knitr::include_graphics("app_figure/workflow_fig.png")
```

@@

#ggsave("A-IAAs_alignment.png", width = 6, height = 4)
```

@@

#ggsave("IAA-tree.png", width = 6, height = 5.5)
```

>>>>

(ref:app-workflow) **ViVa workflow** (a) Workflow diagram of ViVa. Blue indicates user actions, yellow indicates processing steps performed by the application. (b) Detailed look at mapping tab parameters: User selects which genes to look at (1) then clicks the Submit button (2). The "Allele selection" pane is then filled in with all non-reference variants meeting the criteria. The user can adjust the range of nucleotide diversity and the type of SNP with a slider and radio buttons (3) to make the list of variants a manageable size. The list of variants is updated as changes are made to these controls. The user then selects variants to display on the map (4). Clicking the Update Map button (5) populates the map below with points located at the collection coordinates of each accession and colored by the selected alleles. (c) Detailed look at the browser tab options: The user first selects which genes to analyze and clicks submit (1). The “hide 0|0 genotype?” checkbox (2) removes rows from the table containing the reference allele. Four configurable filters of two types are provided. The first filter type is text matching (3); the user selects a column to filter from a drop-down menu then enters one or more text strings to match in that column. The second type of filter is numeric range matching (4); the user again selects a column, then specifies a range to match by typing or selecting minimum and maximum values. After configuring the filters, the user clicks the Apply Filters button (5) to update the table. The table is updated such that only rows that meet all filter conditions are displayed.
```{r app-workflow, fig.cap = "(ref:app-workflow)"}
knitr::include_graphics("app_figure/workflow_fig.png")

```

@@

ggsave("A-IAAs_alignment.tiff", dpi = 800)
```

@@

ggsave("IAA-tree.tiff", dpi = 800)
```

### 2019-09-26 manuscript/06-supplement.Rmd
final submission, minus a few copyedits

#ggsave("IAAs.png", width = 7, height = 4)
```

```{r IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)", fig.height=4}
coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
p <- plotCodingDiv(uniqueCodingVars = coding_variants)
p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, 
             color = "lightgrey", alpha =.1) + 
  annotate("text", x = 116.5, y = 0.1, 
           label = 
             "region of positive selection\nidentified in Winkler et al.") + labs(color = "", y = "nucleotide diversity,\nlog scale") + theme(legend.position = "bottom", legend.text = element_text(size = 8), axis.title = element_text(size = 12)) + guides(color=guide_legend(nrow=5))
```
(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position. The region of positive selection identified by Winkler et al. is highlighted.

@@

```{r AFBtree, fig.cap="(ref:AFBtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% AFB_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio)), 
                                      to = c(0,1)), guide = "colourbar", 
                     breaks = round(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio), 1)
                                    ,2), 
                     labels = round(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio), 1),
                                    2))+
  scale_size(guide = "none") +
  labs(fill = expression(pi [N] / pi [S])) + 
  theme(legend.position = c(.1,.5)) + 
  #geom_nodelab(mapping = aes(label = node)) +
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.2, nudge_y = .2, size = 3) + 
  geom_tiplab(hjust = -0.25) + 
  scale_color_viridis(discrete = TRUE)  
p <- p %>% ggtree::rotate(12) %>% ggtree::rotate(9)
p
#ggsave("AFBs_tree.pdf", height = 3, width = 4)
```
(ref:AFBtree) **AFB protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked with Aliscore [@kuck_parametric_2010] prior to inferring a phylogeny with MrBayes [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$. Nodes are labeled with the posterior probability of monophyly.

@@

@@

```{r TPLtree, fig.cap="(ref:TPLtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% TPL_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio),1), 
                                      to = c(0,1)), guide = "colourbar", 
                     breaks = c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio), 1), 
                     labels = round(c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio), 1),
                                    2)) + scale_size(guide = "none") +
  labs(fill = expression(pi [N] / pi [S])) + 
  theme(legend.position = c(.1,.5)) + 
  #geom_nodelab(mapping = aes(label = node)) +
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.2, nudge_y = .2, size = 3) + 
  geom_tiplab(hjust = -0.25) 
p
```

@@

@@

```{r ARFtree, fig.cap="(ref:ARFtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% 
  ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) +
  geom_tippoint(mapping = 
                  aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio),
                shape = 21, show.legend = TRUE, color = "white",
                position = position_nudge(x = 0.1)) + 
  geom_rootpoint(color = "grey") + 
  geom_nodelab(mapping = aes(label = round(as.numeric(prob), 
                                           digits = 2)), 
               nudge_x = -.3, nudge_y = .5, size = 3) +
  scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 
                                      to = c(0,1)), guide = "colourbar",
                     breaks = c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 
                     labels = round(c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 2)) +
  scale_size(guide = "none") + 
  theme(legend.position = c(.1,.4)) + 
  geom_tiplab(mapping = aes(color = class), hjust = -0.4) + 
  scale_color_viridis(end = .9, discrete = TRUE, option = "A") + 
  labs(fill = expression(pi [N] / pi [S]), color = "ARF class")
p %<>% ggtree::flip(28,32) %>% ggtree::rotate(35)
p 
```

@@

>>>>

ggsave("full-IAA-alignment.tiff", dpi = 800)
```

```{r IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)", fig.height=4}
coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
p <- plotCodingDiv(uniqueCodingVars = coding_variants)
p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, 
             color = "lightgrey", alpha =.1) + 
  annotate("text", x = 116.5, y = 0.1, 
           label = 
             "region of positive selection\nidentified in Winkler et al.") + labs(color = "", y = "nucleotide diversity,\nlog scale") + theme(legend.position = "bottom", legend.text = element_text(size = 8), axis.title = element_text(size = 12)) + guides(color=guide_legend(nrow=5))

ggsave("IAA6-div-plot.tiff", dpi = 800)
```
(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position. The region of positive selection identified by Winkler et al. is highlighted.

@@

```{r AFBtree, fig.cap="(ref:AFBtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% AFB_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio)), 
                                      to = c(0,1)), guide = "colourbar", 
                     breaks = round(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio), 1)
                                    ,2), 
                     labels = round(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio), 1),
                                    2))+
  scale_size(guide = "none") +
  labs(fill = expression(pi [N] / pi [S])) + 
  theme(legend.position = c(.1,.5)) + 
  #geom_nodelab(mapping = aes(label = node)) +
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.2, nudge_y = .2, size = 3) + 
  geom_tiplab(hjust = -0.25) + 
  scale_color_viridis(discrete = TRUE)  
p <- p %>% ggtree::rotate(12) %>% ggtree::rotate(9)
p

ggsave("AFB_tree.tiff", dpi = 800)
```
(ref:AFBtree) **AFB protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked with Aliscore [@kuck_parametric_2010] prior to inferring a phylogeny with MrBayes [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$. Nodes are labeled with the posterior probability of monophyly.

@@

ggsave("full-AFBs-alignment.tiff", dpi = 800)
```

@@

```{r TPLtree, fig.cap="(ref:TPLtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% TPL_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio),1), 
                                      to = c(0,1)), guide = "colourbar", 
                     breaks = c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio), 1), 
                     labels = round(c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio), 1),
                                    2)) + scale_size(guide = "none") +
  labs(fill = expression(pi [N] / pi [S])) + 
  theme(legend.position = c(.1,.5)) + 
  #geom_nodelab(mapping = aes(label = node)) +
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.2, nudge_y = .2, size = 3) + 
  geom_tiplab(hjust = -0.25) 
p

ggsave("TPLtree.tiff", dpi = 800)
```

@@

ggsave("full-TPLs-alignment.tiff", dpi = 800)
```

@@

```{r ARFtree, fig.cap="(ref:ARFtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% 
  ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) +
  geom_tippoint(mapping = 
                  aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio),
                shape = 21, show.legend = TRUE, color = "white",
                position = position_nudge(x = 0.1)) + 
  geom_rootpoint(color = "grey") + 
  geom_nodelab(mapping = aes(label = round(as.numeric(prob), 
                                           digits = 2)), 
               nudge_x = -.3, nudge_y = .5, size = 3) +
  scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 
                                      to = c(0,1)), guide = "colourbar",
                     breaks = c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 
                     labels = round(c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 2)) +
  scale_size(guide = "none") + 
  theme(legend.position = c(.1,.4)) + 
  geom_tiplab(mapping = aes(color = class), hjust = -0.4) + 
  scale_color_viridis(end = .9, discrete = TRUE, option = "A") + 
  labs(fill = expression(pi [N] / pi [S]), color = "ARF class")
p %<>% ggtree::flip(28,32) %>% ggtree::rotate(35)
p

ggsave("ARFtree.tiff", dpi = 800)
```

@@

ggsave("full-ARFs-alignment.tiff", dpi = 800)
